---
layout: page
title: Condition
subtitle: Ogle's Notes
author: Derek H. Ogle
css: /css/codebox.css
csl: ../../rhelpers/american-fisheries-society.csl
bibliography: ../../rhelpers/NRS349.bib
output: 
  html_document: 
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
---
```{r echo=FALSE, eval=FALSE}
# First line renders an appropriate HTML file for the webpage
# Second line makes the script file
# RUN BOTH MANUALLY (following using Knit HTML button)
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Condition/")
source("../../rhelpers/Rhelpers.R")
modHTML("BKG",need2render=FALSE)
if (require(FSA)) purl2("BKG.Rmd",topnotes="")
```

```{r echo=FALSE, results='hide', message=FALSE, warnings=FALSE}
# ===== BEGIN -- THIS CAN BE IGNORED =========================
# Setup of knitr
source("../../rhelpers/knitr_setup.R")
# declare packages used
rqrd <- c("FSA","FSAdata","captioner","knitr","dplyr","magrittr")
library(FSA)
library(FSAdata)
library(dplyr)
library(magrittr)
# setup figure, table, and equation captioning
library(captioner)
eqncaps <- captioner(prefix="Equation")
eqncaps("RelativeWeight")
# ===== END -- THIS CAN BE IGNORED ===========================
```

```{r echo=FALSE}
data(RuffeSLRH92)
ruffe2 <- RuffeSLRH92 %>%
  filterD(!is.na(weight),!is.na(length)) %>%
  mutate(logL=log(length),logW=log(weight))
```

# Body Condition

Condition is a measure of the physical health of a population of fish based on the fish's relative plumpness or fatness. Most often condition is computed by comparing the actual weight of a fish to some expectation of weight based on the length of the fish.  In other words, measuring the condition of a particular fish is an exercise in determining if it weighs more or less than would be expected based on its length.  An overall measure of condition for an entire population is obtained by averaging the condition of all fish in a sample.

The utility of measuring fish conditions was summarized by @Blackwelletal2000 in the following manner:

> Fish condition can be extremely important to fisheries managers.  Plump fish may be indicators of favorable environmental conditions (e.g., habitat conditions, ample prey availability), whereas thin fish may indicate less favorable environmental conditions.  Thus, being able to monitor fish well-being can be extremely useful for fisheries biologists who must make management recommendations concerning fish populations.

There are at least eight metrics of condition [@BolgerConnolly1989] of which three are commonly used by fisheries managers [@Blackwelletal2000].  These three measures are introduced below and discussed within the context of four properties that should be evident in all condition metrics [@Murphyetal1990].  Those four properties are (1) consistency -- similar statistical properties and meaning regardless of species or length; (2) tractability -- analysis by standard statistical methods; (3) efficiency -- relative precision from small samples; and (4) robustness -- relative insensitivity to variations in the way the data was collected and analyzed.

### Fulton's Condition Factor

Fulton's condition factor is calculated with

$$ K = \frac{W}{L^{3}}*constant $$

where the constant is simply a scaling factor that is equal to $100000$ if metric units are used (i.e., grams and millimeters) or $10000$ if English units are used (i.e., pounds and inches).{^1]  Fulton's condition factor *assumes isometric growth*.[^2]  If a fish stock does not exhibit isometric growth, which is often the case, then $K$ tends to differ depending on the length of the fish, violating the consistency property.  Furthermore, comparing $K$ between species is problematic because both species would need to exhibit isometric growth for the comparison to be valid.  Because of these limitations, Fulton's condition factor should be avoided.

### LeCren's Relative Condition Factor

The relative condition factor, introduced by @LeCren1951, is calculated with

$$ Kn = \frac{W}{W'} $$

where $W'$ is the predicted length-specific mean weight for the population under study [@Blackwelletal2000}.  The average $Kn$ across all lengths and species is 1.0 [@AndersonNeumann1996}.  Thus, $Kn$ is consistent across lengths.  @BolgerConnolly1989, however, show that $Kn$ comparisons are restricted to species or regions that have the same slope ($b$) in the length-weight relationship.  Thus, $W'$ is generally predicted from length-weight equations developed for a population (perhaps from several years of data) or for a region.

### Relative Weight

The relative weight, introduced by @WegeAnderson1978, is calculated with

$$ Wr = \frac{W}{Ws}*100 \quad \quad \text{`r paste0("(",eqncaps("RelativeWeight",display="num"),")")`} $$

where $Ws$ is a ``standard weight'' for fish of the same length.  In simplistic terms, a standard weight equation for a particular species is a length-weight relationship designed to predict the 75th percentile[^3] mean weight for a given value of length.[^4]  Standard weight equations have been developed for a wide variety of species.[^5]  It should be noted that the standard weight equations have been developed for either metric or English unit measurements but that the $log_{10}$ rather than the $log_{e}$ transformation is used.  @Blackwelletal2000 suggest that regional or population-specific $Ws$ equations should NOT be developed.  If regional or population-specific analysis is desired then $Kn$ should be used.

Recent trends, following the work of @Gerowetal2005, have resulted in standard weight equations that are quadratic rather than linear.  Examples of the use of quadratic standard weight equations can be found in @OgleWinfield2009 and @CooneyKwak2010.  While use of these types of standard weight equations is not illustrated in this vignette, their use is a simple and straightforward modification of what is demonstrated here.

The relative weight measure has become the most popular measure of condition [@Blackwelletal2000].  This popularity is partly due to the fact that relative weight summaries have been used as a surrogate measure of the general "health" of the fish (@BrownMurphy1991, @NeumannMurphy1992, @Jonasetal1996, @BrownMurphy2004, @Kaufmanetal2007, @RennieVerdon2008}; but also see @Copelandetal2008) as well as the environment (@Liaoetal1995, @Blackwelletal2000, @RennieVerdon2008).  Thus, relative weight summaries may be used as an indirect means for evaluating ecological relationships and the effects of management strategies (@Murphyetal1991, @Blackwelletal2000).  In addition, @Murphyetal1990 found the distributions of $Wr$ values to typically be symmetric (but not normal).  Because t-tests and analysis of variance tests are relatively robust to departures from normality, as long as the distribution is symmetric, typical parametric inferential statistics can be used with $Wr$ values.

It should be noted, though, that Gerow (@Gerowetal2004, @Gerowetal2005, and @Gerow2010) has been critical of the idea that traditionally developed standard weight equations produce relative weight values that are not dependent on the length of the fish.

### Comparisons Among Length Categories

A measure of overall condition using $Wr$ should not be computed without first determining if the $Wr$ values differ across fish lengths [@Blackwelletal2000].  In particular, @Murphyetal1991 suggested that $Wr$ values should first be summarized within the usual five-cell length categories of @Gabelhouse1984.[^6]  Use of the five-cell model can be problematic at times because of small sample sizes in the larger length categories.  Thus, other authors have summarized by 25- or 50-mm length categories.  The mean $Wr$ values in the length categories are then tested with analysis of variance methods to determine if differences exist among the length categories.  Adjacent length categories that are statistically equal can then be pooled together.  If no statistical differences among categories exist, then all length categories can be pooled and an overall measure of condition for the population can be computed.  Relative weights should be reported as whole numbers [@Blackwelletal2000].

\ 

# Calculations in R

Methods for performing these calculations in R are described in Sections 7.1-7.3 and Chapter 8 of @Ogle2016.[^IFARScripts]

\ 

--------------------------------------------------------------

```{r echo=FALSE, results='asis'}
reproInfo(rqrdPkgs=rqrd,out="markdown",links=c(Script="BKG.R",RMarkdown="BKG.Rmd"))
```

--------------------------------------------------------------

\ 

## References

[^IFARScripts]: Scripts for these calculations [are here](http://derekogle.com/IFAR/scripts/).

[^1]: These constants are used so that the values of $K$ are not extremely small fractions.

[^2]: See the length-weight vignette for a brief discussion of isometric and allometric growth.

[^3]: However, see @OgleWinfield2009 for an equation that uses the 50th percentile in addition to the 75th percentile.

[^4]: Specific discussion of methods for computing the standard weight equations are discussed in detail in @Murphyetal1990 and @Blackwelletal2000.

[^5]: In R, examine the \R{data(WSlit)} data frame for all known equations

[^6]: The five-cell length categorization scheme of @Gabelhouse1984 was defined in the size structure module.
