---
layout: page
title: Homework Key
subtitle: Condition
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Condition/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source("../../rhelpers/knitr_setup.R")
options(show.signif.stars=FALSE)
```
```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(FSA)
library(FSAdata)
library(NCStats)
library(dplyr)
library(multcomp)
```
```{r analysis, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
data(InchLake2)
wsv <- wsVal("Bluegill",simplify=TRUE)
psdv <- psdVal("Bluegill")
bg <- filterD(InchLake2,species=="Bluegill") %>%
  mutate(length=length*25.4,
         loglen=log10(length),logwt=log10(weight),
         lcat=lencat(length,breaks=psdv,use.names=TRUE),
         ws=10^(wsv[["int"]]+wsv[["slope"]]*loglen),
         wr=weight/ws*100) %>%
  filterD(length>wsv[["min.TL"]])
```
```{r analysis07, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
bg07 <- filterD(bg,year==2007)
aov07 <- lm(wr~lcat,data=bg07)
anova(aov07)
mc07 <- glht(aov07,mcp(lcat="Tukey"))
summary(mc07)
cld(mc07)
```
```{r analysis08, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
bg08 <- filterD(bg,year==2008)
aov08 <- lm(wr~lcat,data=bg08)
anova(aov08)
mc08 <- glht(aov08,mcp(lcat="Tukey"))
summary(mc08)
cld(mc08)
```

----

## Inch Lake Bluegill

1. 2007
    1. Shown below.
    1. Yes, `r kPvalue(anova(aov07)[1,"Pr(>F)"],latex=FALSE)`
    1. Shown below.
```{r echo=FALSE}
Summarize(wr~lcat,data=bg07,digits=1)
```

```{r echo=FALSE}
fitPlot(aov07,ylim=c(60,100))
addSigLetters(aov07,lets=cld(mc07)$mcletters$Letters,pos=c(2,2,4))
```
    
1. 2008
    1. Shown below.
    1. Yes, `r kPvalue(anova(aov08)[1,"Pr(>F)"],latex=FALSE)`
    1. Shown below.
```{r echo=FALSE}
Summarize(wr~lcat,data=bg08,digits=1)
```

```{r echo=FALSE}
fitPlot(aov08,ylim=c(60,100))
addSigLetters(aov08,lets=cld(mc08)$mcletters$Letters,pos=c(2,2,4))
```

1. Comparison
    1. Smaller fish are skinnier than larger fish -- perhaps more density-dependent competition for smaller fish that they are release from as they get larger.
    1. General decline in condition from 2007 to 2008 -- perhaps poorer growing conditions in 2008, higher density in 2008, or effect of sampling (perhaps different times in each year).

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<setup>>
<<analysis>>
<<analysis07>>
<<analysis08>>
```
