---
layout: page
title: Homework Key
subtitle: Stock-Recruitment
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Recruitment/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source("../../rhelpers/knitr_setup.R")
options(show.signif.stars=FALSE)
```
```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(FSA)
library(FSAdata)
library(nlstools)
library(dplyr)
library(magrittr)

data(WalleyeEL)
WalleyeEL %<>% mutate(logage0=log(age0),logage5=log(age5))
mnS <- mean(WalleyeEL$age5)
rckr <- srFuns(type="Ricker")
ind <- srFuns(type="independence")
```

----

## Escanaba Lake Walleye

```{r RickerMultFit1, echo=FALSE, results='hide', warning=FALSE}
## Multiplicative Model
svR <- srStarts(age0~age5,data=WalleyeEL,type="Ricker")
srR <- nls(logage0~log(rckr(age5,a,b)),data=WalleyeEL,start=svR)
```
```{r RickerMultFit2, echo=FALSE, eval=FALSE}
bootR <- nlsBoot(srR)
```
```{r echo=FALSE, results='hide', warning=FALSE}
load("bootR.Rdata")
```
```{r RickerMultFit3, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
cbind(estimates=coef(srR),confint(srR),confint(bootR))
tmp <- apply(bootR$coefboot,MARGIN=1,FUN=rckr,S=mnS)
predMeanR <- c(Estimate=rckr(mnS,a=coef(srR)),quantile(tmp,c(0.025,0.975)))
```

```{r IndepMultFit, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
## Independence Model
svI <- srStarts(age0~age5,data=WalleyeEL,type="independence")
srI <- nls(logage0~log(ind(age5,a)),data=WalleyeEL,start=svI)
c(estimates=coef(srI),confint(srI))
predMeanI <- ind(mnS,a=coef(srI))
```

```{r Test, echo=FALSE, results='hide'}
( test <- extraSS(srI,com=srR) )
( qrsq <- cor(WalleyeEL$age0,predict(srR))^2)
```

1. The abundance of age-0 fish are the recruits and the abundance of age-5 fish is the spawning stock.
1. The relationship between the abundance of age-0 and age-5 fish is weak at best (see figure further below). I do not expect the Ricker stock-recruitment model to fit well.
1. The equation for the density independent model is R=`r formatC(coef(srI)[["a"]],format="f",digits=2)`S. The equation for the Ricker model is R=`r formatC(coef(srR)[["a"]],format="f",digits=2)`S*exp(-`r formatC(coef(srR)[["b"]],format="f",digits=5)`S).
1. The figure is shown further below.
1. The estimated recruitment at the mean stock level of `r formatC(mnS,format="f",digits=0)` is `r formatC(predMeanI,format="f",digits=0)` for the independence model and `r formatC(predMeanR[["Estimate"]],format="f",digits=0)` (95% CI: `r formatC(predMeanR[["2.5%"]],format="f",digits=0)`,`r formatC(predMeanR[["97.5%"]],format="f",digits=0)`) for the Ricker model.
1. The density-dependent Ricker model fits the data significantly better than then density-independent model (`r kPvalue(test[1,"Pr(>F)"])`). However, the proportion of variability explained is very low (quasi-r<sup>2</sup>=`r formatC(qrsq,format="f",digits=2)`).

```{r fitPlot, echo=FALSE, warning=FALSE}
x <- seq(0,3000,length.out=499)     # many S for prediction
pRR <- rckr(x,a=coef(srR))          # predicted mean R for Ricker
pIR <- ind(x,a=coef(srI))           # predicted mean R for independence model
LCIR <- UCIR <- LCII <- UCII <- numeric(length(x))

for(i in 1:length(x)) {             # CIs for mean R @ each S
  tmp <- apply(bootR$coefboot,MARGIN=1,FUN=rckr,S=x[i])
  LCIR[i] <- quantile(tmp,0.025)
  UCIR[i] <- quantile(tmp,0.975)
}
ylmts <- range(c(pRR,LCIR,UCIR,pIR,LCII,UCII,WalleyeEL$age0))
xlmts <- range(c(x,WalleyeEL$age5))

plot(age0~age5,data=WalleyeEL,xlim=xlmts,ylim=ylmts,col="white",
     ylab="Age-0 Walleye",xlab="Age-5 Walleye")
polygon(c(x,rev(x)),c(LCIR,rev(UCIR)),col=rgb(0,0,0,1/3),border=NA)

points(age0~age5,data=WalleyeEL,pch=19)
lines(pRR~x,lwd=2)
lines(pIR~x,lwd=2,lty=2)
legend("topright",c("Ricker","Independence"),lwd=2,lty=1:2,bty="n")
```

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<setup>>
<<RickerMultFit1>>
<<RickerMultFit2>>
<<RickerMultFit3>>
<<IndepMultFit>>
<<Test>>
<<fitPlot>>
```
