---
layout: page
title: Mark-Recapture II Notes
author: Derek H. Ogle
css: /css/codebox.css
csl: ../../../rhelpers/american-fisheries-society.csl
bibliography: ../../../rhelpers/NRS349.bib
output: 
  html_document: 
    fig_width: 3.5
    fig_height: 3.5
    self_contained: false
    lib_dir: zlibs
---
```{r echo=FALSE, eval=FALSE}
# First line renders an appropriate HTML file for the webpage
# Second line makes the script file
# RUN BOTH MANUALLY (following using Knit HTML button)
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/PREP/NOTES/")
source("../../../rhelpers/Rhelpers.R")
modHTML("MarkRecapture2",need2render=FALSE)
if (require(FSA)) purl2("MarkRecapture2.Rmd",topnotes="")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#### #### #### #### #### #### #### #### #### #### #### #### #### #### #### ##
# ===== BEGIN -- THIS CAN BE IGNORED =========================
# Setup of knitr
source("../../../rhelpers/knitr_setup.R")
# declare packages used
rqrd <- c("FSA","captioner","knitr")
library(FSA)

# ===== END -- THIS CAN BE IGNORED ===========================
#### #### #### #### #### #### #### #### #### #### #### #### #### #### #### ##
```

# Assumptions of Petersen & Related Methods
Before specifically adressing the assumptions of the Petersen and related methods it is important to note that these estimates refer only to the catchable portion or the portion of the population fully recruited to the gear. For example, if the population of northern pike studied was restricted to that part of the population longer than 450 mm because of the gear used to collect the samples, then the estimates of abundance refer only to that part of the population larger than 450 mm. In general, the population estimate only refers to the population extant at the time of the first sample, except under very specific conditions which are outlined below.

The Petersen and related methods depend on meeting five assumptions [@Seber1982]:

* The population is physically (i.e., no immigration or emigration) and demographically (i.e., no recruitment or mortality) closed (so that $N$ is constant over the time of sampling).
* Marks or tags are neither lost nor missed
* Marked individuals returned to the population mix randomly with unmarked individuals.
* All individuals within a sample have an equal probability of capture.
* Individual behavior or vulnerability does not change after being marked.

Fisheries personnel attempt to control the first assumption by having a very short period between the first and second samples. However, there are several ways in which the first assumption can be violated without adversely affecting the usability of the Petersen estimate [@Seber1982]. If fish die while being marked and before being released, then the Petersen estimate will refer to the size of the population **after** the marked fish from the first sample are released. Mortality between the first and second sample will not affect the Petersen estimate of initial population size if the rate of mortality is the **same** for both marked and unmarked individuals (i.e., so the proportion of marked animals in the population is not affected). Recruitment of new individuals to the population will completely invalidate the Petersen method as an estimator of the initial population size. However, if there is no mortality along with the recruitment, then the Petersen method will provide a valid estimate of the population size at the time of the **second** sample. @Seber1982 [p.73] provides a statistical test to detect recruitment between the two samples.

The second assumption is difficult to assure because it is impossible to take a strict random sample in most real situations. If all fish are equally catchable then a random sample can be approximated by sampling areas at random with constant effort [@Krebs1999]. Systematic sampling is often used to select animals but this relies on the assumption that marked and unmarked animals mix uniformly through the population.

Unequal catchability between marked and unmarked fish may arise from three general causes [@Eberhardt1969]:

* The behavior of individuals in the vicinity of the trap.
* Learning by animals already caught (i.e., trap-shy or trap-happiness).
* Unequal opportunity to be caught because of trap positions.

Statistical tests exist (see @Krebs1999) to identify certain forms of unequal catchability, but all of these tests rely on more than two samples from the population. Thus, in a Petersen type study, the biologist must take care to design fish collections that minimize these difficulties.

Finally, violations of the last two assumptions result in the possible underestimation of the proportion of marked animals in the second sample which leads to overestimation of the population size. The fourth assumption can be tested by "double-marking" a portion of the marked animals. For example, the northern pike in Harding Lake were marked with a numbered floy tag and a partial fin-clip. The proportion of lost-tags can be estimated from the recaptured fish and a correction applied to the population estimate.

\ 

--------------------------------------------------------------

```{r echo=FALSE, results='asis'}
reproInfo(rqrdPkgs=rqrd,out="markdown",
          links=c(Script="MarkRecapture2.R",RMarkdown="MarkRecapture2.Rmd"))
```

--------------------------------------------------------------

\ 

## References
