---
layout: page
title: Homework Key
subtitle: Growth
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Growth/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source("../../rhelpers/knitr_setup.R")
options(show.signif.stars=FALSE)
```
```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(FSA)
library(FSAdata)
library(nlstools)
library(dplyr)
library(magrittr)

data(SculpinALTER)
vbTyp <- vbFuns("Typical")
```

----

## Alaskan Slimy Sculpin A

```{r TypicalFit1, echo=FALSE, results='hide', warning=FALSE}
svTyp <- vbStarts(tl~age,data=SculpinALTER,type="Typical")
fitTyp <- nls(tl~vbTyp(age,Linf,K,t0),data=SculpinALTER,start=svTyp)
```
```{r TypicalFit2, eval=FALSE, echo=FALSE, results='hide', warning=FALSE}
bootTyp <- nlsBoot(fitTyp)
```
```{r echo=FALSE, results='hide', warning=FALSE}
load("bootTyp.Rdata")
```
```{r TypicalFit3, echo=FALSE, results='hide', warning=FALSE}
estsTyp <- cbind(Est=coef(fitTyp),confint(bootTyp))
p3Typ <- apply(bootTyp$coefboot,MARGIN=1,FUN=vbTyp,t=3)
p3Typ <- c(Est=vbTyp(3,coef(fitTyp)),quantile(p3Typ,c(0.025,0.975)))
halfLinf <- log(2)/coef(fitTyp)[["K"]]
```

1. The plot is shown below. The data show a fairly obvious curve, a fairly obious asymptote, no fish less than age-1 are shown, and there is considerable variable in length at each age.
1. The plot is shown below.
1. The residual plot is shown below. There is a slight heteroscedasticity, no nonlinear, and approximate normality of the residuals.
1. The typical von Bertalanffy equation is $E(L|t)=`r formatC(estsTyp["Linf","Est"],format="f",digits=1)`(1-e^{-`r formatC(estsTyp["K","Est"],format="f",digits=2)`(t+`r formatC(-1*estsTyp["t0","Est"],format="f",digits=2)`)})$.
1. The asymptotic mean length of Slimy Sculpins is `r formatC(estsTyp["Linf","Est"],format="f",digits=1)`. The Brody growth coefficent, or the rate at which the mean length approaches the asymptotic mean length, is `r formatC(estsTyp["K","Est"],format="f",digits=2)`. The age at which the mean length is zero (i.e,. the x-intercept) is `r formatC(estsTyp["t0","Est"],format="f",digits=2)`.
1. The 95% confidence interval for the asymptotic mean length is `r formatC(estsTyp["Linf","95% LCI"],format="f",digits=1)` to `r formatC(estsTyp["Linf","95% UCI"],format="f",digits=1)`, for the Brody growth coefficient is `r formatC(estsTyp["K","95% LCI"],format="f",digits=2)` to `r formatC(estsTyp["K","95% UCI"],format="f",digits=2)`, and for the x-intercept is `r formatC(estsTyp["t0","95% LCI"],format="f",digits=2)` to `r formatC(estsTyp["t0","95% UCI"],format="f",digits=2)`. The intervals for the asymptotic mean length and Brody growth coefficient are quite wide.
1. The predicted mean length at age-3 is `r formatC(p3Typ["Est"],format="f",digits=1)` (with 95% confidence interval from `r formatC(p3Typ["2.5%"],format="f",digits=1)` to `r formatC(p3Typ["97.5%"],format="f",digits=1)`). This interval is fairly narrow because it is a confidence (rather than prediction) interval and age-3 is a well-represented age in the data.
1. Slimy Sculpins will have reached half of their asymptotic length at age `r formatC(halfLinf,format="f",digits=2)`.

```{r fitPlotA, echo=FALSE}
x <- seq(0,5,length.out=199)
pTyp <- vbTyp(x,Linf=coef(fitTyp))
xlmts <- range(c(x,SculpinALTER$age))
ylmts <- range(c(pTyp,SculpinALTER$tl))
plot(tl~age,data=SculpinALTER,xlab="Age", ylab="Total Length (mm)",xlim=xlmts,
     ylim=ylmts,pch=19,col=rgb(0,0,0,1/3))
lines(pTyp~x,lwd=2)
LCI <- UCI <- numeric(length(x))
for(i in 1:length(x)) {
  tmp <- apply(bootTyp$coefboot,MARGIN=1,FUN=vbTyp,t=x[i]) 
  LCI[i] <- quantile(tmp,0.025) 
  UCI[i] <- quantile(tmp,0.975)
}
ylmts <- range(c(pTyp,LCI,UCI,SculpinALTER$tl))
lines(UCI~x,lwd=2,lty="dashed")
lines(LCI~x,lwd=2,lty="dashed")
```
```{r residPlotA, echo=FALSE}
residPlot(fitTyp)
```


#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<setup>>
<<TypicalFit1>>
<<TypicalFit2>>
<<TypicalFit3>>
<<fitPlotA>>
<<residPlotA>>
```

## Alaskan Slimy Sculpin B

```{r OriginalFit1, echo=FALSE, results='hide', warning=FALSE}
vbOrig <- vbFuns("Original")
svOrig <- vbStarts(tl~age,data=SculpinALTER,type="Original")
fitOrig <- nls(tl~vbOrig(age,Linf,K,L0),data=SculpinALTER,start=svOrig)
```
```{r OriginalFit2, eval=FALSE, echo=FALSE, results='hide', warning=FALSE}
bootOrig <- nlsBoot(fitOrig)
```
```{r echo=FALSE, results='hide', warning=FALSE}
load("bootOrig.Rdata")
```
```{r OriginalFit3, echo=FALSE, results='hide', warning=FALSE}
estsOrig <- cbind(Est=coef(fitOrig),confint(bootOrig))
p3Orig <- apply(bootOrig$coefboot,MARGIN=1,FUN=vbOrig,t=3)
p3Orig <- c(Est=vbOrig(3,coef(fitOrig)),quantile(p3Orig,c(0.025,0.975)))
```

1. The plot is shown below.
1. The residual plot is shown below. There is a slight heteroscedasticity, no nonlinear, and approximate normality of the residuals.
1. The typical von Bertalanffy equation is $E(L|t)=`r formatC(estsOrig["Linf","Est"],format="f",digits=1)`-(`r formatC(estsOrig["Linf","Est"],format="f",digits=1)`-`r formatC(estsOrig["L0","Est"],format="f",digits=1)`)(1-e^{-`r formatC(estsOrig["K","Est"],format="f",digits=2)`t})$.
1. The asymptotic mean length of Slimy Sculpins is `r formatC(estsOrig["Linf","Est"],format="f",digits=1)`. The Brody growth coefficent, or the rate at which the mean length approaches the asymptotic mean length, is `r formatC(estsOrig["K","Est"],format="f",digits=2)`. The mean length at age-0 is (i.e., the y-intercept) is `r formatC(estsOrig["L0","Est"],format="f",digits=1)`.
1. The common parameter estimates from the typical and the original parameterizations of the von Bertalanffy are the same. This is because a parameterization is the same model, just with different parameters.
1. The predicted mean length at age-3 is `r formatC(p3Orig["Est"],format="f",digits=1)` (with 95% confidence interval from `r formatC(p3Orig["2.5%"],format="f",digits=1)` to `r formatC(p3Orig["97.5%"],format="f",digits=1)`). This prediction is the same as that from fitting the typical model (within rounding for the confidence intervals).

```{r fitPlotB, echo=FALSE}
x <- seq(0,5,length.out=199)
pOrig <- vbOrig(x,Linf=coef(fitOrig))
xlmts <- range(c(x,SculpinALTER$age))
ylmts <- range(c(pTyp,SculpinALTER$tl))
plot(tl~age,data=SculpinALTER,xlab="Age", ylab="Total Length (mm)",xlim=xlmts,
     ylim=ylmts,pch=19,col=rgb(0,0,0,1/3))
lines(pOrig~x,lwd=2)
LCI <- UCI <- numeric(length(x))
for(i in 1:length(x)) {
  tmp <- apply(bootOrig$coefboot,MARGIN=1,FUN=vbOrig,t=x[i]) 
  LCI[i] <- quantile(tmp,0.025) 
  UCI[i] <- quantile(tmp,0.975)
}
ylmts <- range(c(pOrig,LCI,UCI,SculpinALTER$tl))
lines(UCI~x,lwd=2,lty="dashed")
lines(LCI~x,lwd=2,lty="dashed")
```

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<OriginalFit1>>
<<OriginalFit2>>
<<OriginalFit3>>
<<fitPlotB>>
```

