---
layout: page
title: Homework Key
subtitle: Size Structure
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/SizeStructure/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source("../../rhelpers/knitr_setup.R")
options(show.signif.stars=FALSE)
```
```{r setup, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(FSA)
library(FSAdata)
library(dplyr)
library(magrittr)
```
```{r analysis, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
data(YPerchSB1)
cuts <- psdVal("Yellow Perch",units="cm")
YPerchSB1 %<>% mutate(gcat=lencat(tl,breaks=cuts,
                                  drop.levels=TRUE,use.names=TRUE))
ypIn <- filterD(YPerchSB1,loc=="inner")
ypOut <- filterD(YPerchSB1,loc=="outer")

psdIn <- psdCalc(~tl,data=ypIn,species="Yellow Perch",
                 units="cm",what="traditional")
psdOut <- psdCalc(~tl,data=ypOut,species="Yellow Perch",
                  units="cm",what="traditional")

gblhTbl <- xtabs(~loc+gcat,data=YPerchSB1)[,-1]
psdQTbl <- gblhTbl[,-3]
psdQTbl[,"quality"] <- rowSums(gblhTbl[,2:3])
psdPTbl <- gblhTbl[,-2]
psdPTbl[,"stock"] <- rowSums(gblhTbl[,1:2])

ks <- ksTest(tl~loc,data=YPerchSB1)
gblhChi <- chisq.test(gblhTbl)
psdQChi <- chisq.test(psdQTbl)
psdPChi <- chisq.test(psdPTbl)
```

----

## Saginaw Bay Yellow Perch

1. Inner location
    1. Most of the individuals are between 10 and 23 cm. There are no distinctive age-classes.
    1. The percentage of stock-size fish that are also quality-size is 12% with a 95% confidence interval from 10 to 15%.
    1. The percentage of stock-size fish that are also preferred-size is 1% with a 95% confidence interval from 0 to 2%
```{r echo=FALSE}
hist(~tl,data=ypIn,w=1,xlab="Total Length (cm)",xlim=c(5,30))
psdIn
```
    
1. Outer location
    1. Most fish are between about 8 and 30 cm, with a distinct mode between 15 and 20 cm. There is a much smaller mode etween 8 and 14 cm. Fish larger than 20 cm tail off nearly linearly.
    1. The percentage of stock-size fish that are also quality-size is 35% with a 95% confidence interval from 30 to 39%.
    1. The percentage of stock-size fish that are also preferred-size is 11% with a 95% confidence interval from 8 to 14%
```{r echo=FALSE}
hist(~tl,data=ypOut,w=1,xlab="Total Length (cm)",xlim=c(5,30))
psdOut
```

1. Comparison between the inner and outer locations.
    1. There is a significant difference in the length frequency distributions of fish between the inner and outer locations (`r kPvalue(ks$p.value,latex=FALSE)`).
    1. There is a significant difference in the distribution of fish into the Gabelhouse length categories between the inner and outer locations (`r kPvalue(gblhChi$p.value,latex=FALSE)`).
    1. There is a significant difference in PSD-Q values between the inner and outer locations (`r kPvalue(psdQChi$p.value,latex=FALSE)`).
    1. There is a significant difference in PSD-P values between the inner and outer locations (`r kPvalue(psdPChi$p.value,latex=FALSE)`).


#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<setup>>
<<analysis>>
```
