---
layout: page
title: Total Mortality
subtitle: Ogle's Notes
author: Derek H. Ogle
css: /css/codebox.css
csl: ../../rhelpers/american-fisheries-society.csl
bibliography: ../../rhelpers/NRS349.bib
output: 
  html_document: 
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
---
```{r echo=FALSE, eval=FALSE}
# First line renders an appropriate HTML file for the webpage
# Second line makes the script file
# RUN BOTH MANUALLY (following using Knit HTML button)
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Mortality/")
source("../../rhelpers/Rhelpers.R")
modHTML("BKG2",need2render=FALSE)
if (require(FSA)) purl2("BKG.Rmd",topnotes="")
```

```{r echo=FALSE, results='hide', message=FALSE, warnings=FALSE}
# ===== BEGIN -- THIS CAN BE IGNORED =========================
# Setup of knitr
source("../../rhelpers/knitr_setup.R")
# declare packages used
rqrd <- c("FSA","captioner","knitr")
library(FSA)
# setup figure, table, and equation captioning
library(captioner)

tblcaps <- captioner(prefix="Table")
tblcaps("BrkTrtDataCR","Cross-sectional total catch-at-age of Tobin Harbor Brook Trout in fyke nets, 1996-1998, modified to illustrate the calculations of the Chapman-Robson method.")

eqncaps <- captioner(prefix="Equation")
eqncaps("ChapmanRobsonS")
eqncaps("ChapmanRobsonSSE1")
eqncaps("ChapmanRobsonSSE2")
eqncaps("ChapmanRobsonZSE")
eqncaps("ChapmanRobsonZ2")
eqncaps("ChapmanRobsonZSE2")
eqncaps("")
# ===== END -- THIS CAN BE IGNORED ===========================
```

# Chapman-Robson Method
### Background

@ChapmanRobson1960 (and @RobsonChapman1961) provided an alternative method for estimating the total annual survival rate ($S$), and thus the annual ($A$) and instantaneous ($Z$) total mortality rates, from catch curve data.  Their method was based on understanding that the catches at each age on the descending limb of the catch curve followed a geometric probability distribution and using this to derive a maximum likelihood estimator for the survival parameter of the distribution.  Their method, called the Chapman-Robson method, is outlined below.

The Chapman-Robson estimate of the annual survival rate is

$$ \hat{S} = \frac{T}{n+T-1} = \frac{\bar{T}}{1+\bar{T}-\frac{1}{n}} \quad \quad \text{`r paste0("(",eqncaps("ChapmanRobsonS",display="num"),")")`} $$

where $n$ is the total number of fish observed on the descending limb of the catch curve, $T$ is the total recoded age of fish on the descending limb of the catch curve, and $\bar{T}$ is the mean recoded age of fish on the descending limb of the catch curve (i.e., $\bar{T}=\frac{T}{n}$).  It should be noted that the ages are "recoded" such that the first fully-recruited age on the descending limb of the catch-curve is set to 0.  The total recoded age is calculated as a weighted sum of the recoded ages where the weights are the catches at each age.  The standard error of this estimate is

$$ SE_{\hat{S}} = \sqrt{\frac{T}{n+T-1}\left(\frac{T}{n+T-1}-\frac{T-1}{n+T-2}\right)} = \sqrt{\hat{S}\left(\hat{S}-\frac{T-1}{n+T-2}\right)} \quad \quad \text{`r paste0("(",eqncaps("ChapmanRobsonSSE1",display="num"),")")`} $$

If $n$ is large, as it often is in fisheries catch data, then $SE_{\hat{S}}$ can be estimated by

$$ SE_{\hat{S}} = \sqrt{\frac{\hat{S}(1-\hat{S})^{2}}{n}} \quad \quad \text{`r paste0("(",eqncaps("ChapmanRobsonSSE2",display="num"),")")`} $$

The Chapman-Robson estimate of $S$ can be transformed into an estimate of $Z$ through the relationship $S=e^{-Z}$, i.e., $\hat{Z} = -log(\hat{S})$ with a large sample approximation of $SE_{\hat{Z}}$ [@Jensen1985] as

$$ SE_{\hat{Z}} = \frac{SE_{\hat{S}}}{\hat{S}} \quad \quad \text{`r paste0("(",eqncaps("ChapmanRobsonZSE",display="num"),")")`} $$

@Hoenigetal1983, however, have shown that these estimates are slightly biased and that an unbiased estimate of $Z$ is obtained with

$$ \hat{Z} = -log(\hat{S}) - \frac{(n-1)(n-2)}{n(T+1)(N+T-1)} \quad \quad \text{`r paste0("(",eqncaps("ChapmanRobsonZ2",display="num"),")")`} $$

with the large sample approximation of $SE_{\hat{Z}}$ as

$$ SE_{\hat{Z}} = \frac{1-e^{-\hat{Z}}}{\sqrt{ne^{-\hat{Z}}}} \quad \quad \text{`r paste0("(",eqncaps("ChapmanRobsonZSE2",display="num"),")")`} $$

These calculations are illustrated below with the Tobin Harbor Brook Trout (assuming that the fish were fully recruited to the fyke nets at age-2).  The original data were modified for calculation of the Chapman-Robson estimator of $S$ as shown in `r tblcaps("BrkTrtDataCR",display="cite")`.  From that, it is seen that $n=235$ and $T=229$.  Thus, 

$$ \hat{S} = \frac{229}{235+229-1} = 0.4946004 $$

$$ SE_{\hat{S}} = \sqrt{0.4946004\left(0.4946004-\frac{229-1}{235+229-2}\right)} = 0.02326041 $$

An approximate 95% confidence interval for $S$ is $0.4946\pm1.96(0.0233)$ or (0.4490,0.5402).  Furthermore,

$$ \hat{Z} = -log(0.4946004) - \frac{(235-1)(235-2)}{235(229+1)(235+229-1)} = 0.7018264 $$

$$ SE_{\hat{Z}} = \frac{1-e^{-0.7018264}}{\sqrt{235*e^{-0.7018264}}} = 0.04672751 $$

An approximate 95% confidence interval for $Z$ is $0.7018\pm1.96(0.0467)$ or (0.6102,0.7934).

\ 

`r tblcaps("BrkTrtDataCR")`

<table style="width:40%">
<tr><td><b>Age</b></td><td><b>Recoded Age</b></td><td><b>Catch</b></td><td><b>Recode*Catch</b></td></tr>
<tr><td>2</td><td>0</td><td>112</td><td>0</td></tr>
<tr><td>3</td><td>1</td><td>45</td><td>45</td></tr>
<tr><td>4</td><td>2</td><td>58</td><td>116</td></tr>
<tr><td>5</td><td>3</td><td>12</td><td>36</td></tr>
<tr><td>6</td><td>4</td><td>8</td><td>32</td></tr>
<tr><td><b>sum</b></td><td></td><td><b>235</b></td><td><b>229</b></td></tr>
<table>

### Which Ages?

@Smithetal2012 suggest that the Chapman-Robson method should use all ages after the age where the peak catch occurred (i.e., all ages beginning with the first age after the age with the peak catch).

\ 


# Catch Curve vs Chapman-Robson

@Dunnetal2002 provided an excellent review of past examinations of the regression and Chapman-Robson methods and their own examination of the precision and bias properties of these two methods in the face of stochastic errors related to $Z$, number of fish at time of recruitment to the fishery, sampling, and ageing.  Overall, they found that the Chapman-Robson estimator was most precise and least biased; however, the advantage over the regression method declined somewhat with increasing amounts of stochastic error and increasing values of $Z$.

The work of @Dunnetal2002 also showed that, in the face of only stochastic sampling variability, the Chapman-Robson estimator was very slightly positively biased, primarily for larger values of $Z$, but only on the order of approximately 2-3%.  In contrast, the regression estimator had a strong negative bias on the order of 20%.  The modified (excluding all age-classes beyond where one or fewer individuals were observed) regression estimator had a negative bias on the order of 2-5% with the larger values occurring when $Z$ was larger.  These results suggest that estimates of $Z$ with the regression method may be serious underestimates.

@Smithetal2012 provided another study of precision between the regression and Chapman-Robson methods with the addition of consideration of different definitions of the descending limb of the catch curve.  They found that the Chapman-Robson method using all ages after the age with the peak catch and the weighted regression using all ages after and including the age with the peak catch performed similarly.  However, they suggest using the Chapman-Robson method because it is based on a statistical foundation and has a generally smaller variance, whereas the weighting procedure in the regression method is *ad hoc*.  Finally, @Smithetal2012 conclude that the unweighted regression should not be used.

Finally, other methods for estimating total mortality rates have been proposed (e.g., @Heincke1913, @Jackson1939, @SsentengonoLarkin1973).  However, various studies (including @Smithetal2012) have shown that these methods perform less well than the Chapman-Robson and regression methods described here.


# Calculations in R

Methods for performing these calculations in R are described in Sections 11.1 and 11.2 of @Ogle2016.[^IFARScripts]

\ 

--------------------------------------------------------------

```{r echo=FALSE, results='asis'}
reproInfo(rqrdPkgs=rqrd,out="markdown",links=c(Script="BKG2.R",RMarkdown="BKG2.Rmd"))
```

--------------------------------------------------------------

\ 

## References

[^IFARScripts]: Scripts for these calculations [are here](http://derekogle.com/IFAR/scripts/).
