---
layout: page
title: Homework Key
subtitle: Mark-Recapture I
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_KEYS/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_MarkRecapture1_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
source("../../../rhelpers/knitr_setup.R")
```

----

## Ugly River Brown Trout
```{r UglyMR, echo=FALSE, results='hide', message=FALSE}
library(FSA)
mr1 <- mrClosed(43,123,16,method="Chapman")
cbind(summary(mr1),confint(mr1,verbose=TRUE))
```
```{r echo=FALSE, results='hide'}
ci <- confint(mr1)
```

1. M=43, n=123, m=16, N is unknown
1. The estimated abundance of Brown Trout (*Salmo trutta*) in this section of the Ugly River is `r formatC(mr1$N,format="f",digits=0)` fish with a 95% confidence interval from `r formatC(ci[1,"95% LCI"],format="f",digits=0)` to `r formatC(ci[1,"95% UCI"],format="f",digits=0)`.
1. The estimate is not the same as the published result because they did not use the Chapman modification. The confidence interval is very different (the authors used the Poisson approximation rather than the binomial approximation).
1. The binomial distribution was used to construct the confidence intervals because more than 10% of the fish in the second sample were recaptured (i.e., `r formatC(mr1$m/mr1$n*100,format="f",digits=1)`).

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<UglyMR>>
```

----

## Upper Niagara Springs Pond Rainbow Trout
```{r echo=FALSE, results='hide', message=FALSE}
data(RBTroutUNSP,package="FSAdata")
```

```{r UNSPMR, echo=FALSE, results='hide', message=FALSE}
ch1 <- capHistSum(RBTroutUNSP)
mr2 <- mrClosed(ch1,method="Chapman")
cbind(summary(mr2),confint(mr2,verbose=TRUE))
```
```{r echo=FALSE, results='hide'}
ci2 <- confint(mr2)
```

1. A total of `r ch1$caphist[["10"]]` fish were captured in the first but not the second sample, `r ch1$caphist[["01"]]` fish were captured in the second but not the first sample, and `r ch1$caphist[["11"]]` fish were captured in both samples.
1. M=`r ch1$sum[["M"]]`, n=`r ch1$sum[["n"]]`, m=`r ch1$sum[["m"]]`
1. The estimated abundance of Rainbow Trout (*Oncorhynchus mykss*) greater than 100 mm in Upper Niagara Springs Pond is `r formatC(mr2$N,format="f",digits=0)` fish with a 95% confidence interval from `r formatC(ci2[1,"95% LCI"],format="f",digits=0)` to `r formatC(ci2[1,"95% UCI"],format="f",digits=0)`.
1. The estimate is the same as the published result, but the confidence interval is dramatically different (they likely used a normal approximation).
1. The Poisson distribution was used to construct the confidence interval because `r mr2$m` (<50) fish, but not greater than 10% of the fish (i.e., `r formatC(mr2$m/mr2$n*100,format="f",digits=1)`), in the second sample were recaptured.

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
RBTroutUNSP <- read.csv("RBTroutUNSP.csv")
<<UNSPMR>>
```
