---
layout: page
title: Homework Key
subtitle: Mark-Recapture II
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_KEYS/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_MarkRecapture2_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
source("../../../rhelpers/knitr_setup.R")
```

## Observations

1. The initial N can be reliably estimated only if the assumption violation is that there is mortality on both marked and unmarked fish.
1. The final N can be reliably estimated when the assumption violations are mortality on unmarked fish only or recruitment into the population.
1. There is no assumption violation that allows both the initial and final populations to be reliably estimated.
1. The following things should be done to minimize assumption violations:
    a. The marking and recapture run should be relatively close in time to minimize mortality, recruitment, emigration, and immigration.
    a. Technicians must pay attention so that all marked fish that are captured are observed and so that marks/tags are properly affixed.
    a. Steps should be taken to assure that the marked fish mix into the population as a whole (so don't release fish next to the net).
    a. Marks/Tags and the handling of fish should have as little affect on the behavior and survival of the fish as possible (i.e., handle them quickly and carefully).

```{r echo=FALSE, message=FALSE}
tibble::tribble(
  ~Situation,~N_Initial,~N_Final,
  "Lose mark","Over","Over",
  "Mortality (same)","OK","Over",
  "Mortality (Marked)","Over","Over",
  "Mortality (Unmarked)","Under","OK",
  "Recruitment","Over","OK",
  "Trap Happy","Under","Under",
  "Trap Shy","Over","Over",
  "Missed Mark","Over","Over",
  "Released Near Net","Under","Under"
)
```
