---
layout: page
title: Homework Key
subtitle: Course Intro
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_KEYS/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_CourseIntro_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
source("../../../rhelpers/knitr_setup.R")
```

## Lake Mille Lacs Walleye
The Minnesota DNR convened a blue-ribbon panel to determine whether the data available from Lake Mille Lacs supported a variety of hypotheses about why the fish population declined. The report from the blue-ribbon panel is available [here in long form](MilleLacs_Report.pdf) or [here as a summarized set of presentation slides](MilleLacs_Report_Presentation.pdf). Use these resources to answer the questions below.

1. How did the regulations managing the harvest of Walleye (*Sander vitreus*) change around 2000?
    * Establishment of a joint (recreational and tribal) Walleye fishery in 1998,
    * Introduction of harvestable slot limits to the recreational fishery in 1999, and
    * A switch to protected slot limits in 2003.
1. What were two abiotic and three biotic factors that may affect Walleye populations that changed dramatically in the late 1990s or early to mid 2000s?
    * Abiotic: (i) mean annual temperature in northern Minnesota has increased ~1.5<sup>o</sup>C since 1980 and (ii) water clarity has increased 1.5-fold since the mid-1990s.
    * Biotic: (i) Potential Walleye competitors and predators have increased since the mid- to late-1990s, (ii) important Walleye prey have decreased, (iii) invasive Zebra mussels (*Dreissena polymorpha*) and spiny water fleas (*Bythotrephes longimanus*) were first detected in 2005 and 2009, respectively, and (iv) zooplankton biomass has been low since 2012.
1. What were the four working hypotheses that the blue-ribbon panel addressed?
    * Too few spawners, possibly as a result of overharvest,
    * low egg production and/or survival to first fall,
    * low survival from first to second fall, and 
    * low survival from age-1 to fishable ages.
1. For each working hypothesis, summarize the panel's findings, both with some details and with their overall conclusion about the probability that that hypothesis contributed to the decline of the Lake Mille Lacs Walleye population. Your summary should at least comment on their findings for each sub-hypothesis.
    * Too few spawners, possibly as a result of overharvest,
        * Probability: LOW
        * no evidence that adult biomass reached particularly low levels prior to the substantial decline in year class strength
        * no evidence that the decline in the Mille Lacs Walleye population was due to a skewed sex ratio in the adult population.
        * no evidence of a sustained decline in the relative abundance of older or larger adults or the average age of adults
        * fishing and total mortality rates experienced by Mille Lacs Walleye were not excessive on average or for extended periods of time
    * low egg production and/or survival to first fall,
        * Probability: LOW
        * No evidence that the decline in Mille Lacs Walleye was due to low egg production or survival to the first fall of life
    * low survival from first to second fall, and 
        * Probability: HIGH
        * inconclusive evidence that overwinter survival was low for small, age-0 Walleye.
        * Weight data did not support the hypothesis that age-0 Walleye have low energy reserves heading into winter.
        * Walleye cannibalism and perhaps predation by northern pike could be responsible for low overwinter survival of age-0 Walleye in Mille Lacs.
        * Cannibalism and perhaps predation by northern pike and double-crested cormorants could be responsible for low summer survival of age-1 Walleye
    * low survival from age-1 to fishable ages.
        * Probability: High
        * No evidence that Mille Lacs Walleye are food-limited.
        * cannibalism and perhaps predation by northern pike and double-crested cormorants could be responsible for the low survival of age 1-3 Walleye in Mille Lacs
1. What recommendations did the panel make for data collection?
    * Identify and monitor key indicators related to early survival, predator abundance, diet (and consumption), invasive species, and water quality.
    * Characterize pre-2005 Walleye diets (with stable isotope analysis of scales).
    * Sample/analyze Walleye, northern pike, and smallmouth bass diets in both summer and winter
    * Estimate the current and future impact of cormorants on Walleye.
    * Determine why young Walleye survival is low despite high yellow perch abundance.
    * Determine the extent to which body length is contributing to low survival in the first winter of life.
    * Estimate age-specific natural mortality rates.
    * Evaluate how changes to the stock-recruitment relationship, consistent with changes in early-life Walleye survival, influence sustainable fishing
    * Evaluate gear biases over time.
1. What recommendations did the panel make for fisheries management?
    * Manage fishing mortality.
    * Avoid stocking.
