---
layout: page
title: Homework Key
subtitle: Age-Length Key
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Age/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(FSA)
library(magrittr)
library(dplyr)
source("../../rhelpers/knitr_setup.R")
```

----

## Lake Ontario Rock Bass A

```{r LORB, echo=FALSE, results="hide", fig.show="hide"}
df <- read.csv("https://raw.githubusercontent.com/droglenc/FSAdata/master/data-raw/RockBassLO2.csv") %>%
  mutate(lcat10=lencat(tl,w=10))
aged <- filterD(df,!is.na(age))
unaged <- filterD(df,is.na(age))

aged.ld <- xtabs(~lcat10,data=aged)
aged.ad <- xtabs(~age,data=aged)

alk.raw <- xtabs(~lcat10+age,data=aged)
alk <- prop.table(alk.raw,margin=1)

alkPlot(alk)

unaged <- alkIndivAge(alk,age~lcat10,data=unaged)
df2 <- rbind(aged,unaged)

all.ld <- xtabs(~lcat10,data=df2)
all.ad <- xtabs(~age,data=df2)

hist(~age,data=df2,w=1,xlab="Age (yrs)")

all.ls <- group_by(df2,age) %>%
  summarize(mnlen=mean(tl)) %>%
  as.data.frame()

plot(tl~age,data=df2,ylab="Total Length (mm)",xlab="Age (yrs)",
     pch=19,col=col2rgbt("black",1/20))
lines(mnlen~age,data=all.ls,col="blue",lwd=2)
```

1. There are `r nrow(aged)` fish that were aged and `r nrow(unaged)` fish that were not aged.
1. xxx
    1. There were `r aged.ld[["180"]]` fish in the 180-mm category in the aged sample.
    1. There were `r aged.ad[["7"]]` age-7 fish in the aged sample.
    1. The proportion of fish in the 140-mm TL category that were age-4 was `r alk["140","4"]`.
    1. The proportion of fish in the 200-mm TL category that were age-8 was `r alk["200","8"]`.
1. xxx
    1. The proportion of fish in the 210-mm TL category that should be assigned age-5 was `r alk["210","5"]`.
    1. Of 30 Rock Bass in the 180-mm TL category, `r 30*alk["180","5"]` should be assigned age-5.
    1. This age-length key is pretty "clean." There are some anomolies with age-8 and age-9 fish near 200 mm and age-11 fish from 230 to 270 mm, but these proportions are fairly small.
1. Did not do.

## Lake Ontario Rock Bass B
1. Completed above.
1. xxx
    1. There were `r all.ad[["5"]]` age-5 fish after applying the age-length key.
    1. There were `r all.ad[["11"]]` age-11 fish after applying the age-length key.
    1. xxx
    1. There were `r all.ld[["150"]]` total fish in the 150-mm TL category.
    1. The mean TL of age-5 fish is `r formatC(all.ls$mnlen[all.ls$age==5],format="f",digits=1)` mm.
    1. xxx
1. Should not have gotten the exact same results, but they should be close. They are different because of the randomization, but they are similar because little is randomized.

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<LORB>>
```

