---
layout: page
title: Homework Key
subtitle: Mortality
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/Mortality/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(FSA)
library(magrittr)
library(dplyr)
source("../../rhelpers/knitr_setup.R")
```

----

## Keewenaw Bay Lake Trout

```{r KBLT, echo=FALSE, results="hide", fig.show="hide"}
d <- data.frame(age=3:14,ct=c(5,18,21,10,45,109,95,63,42,25,13,4)) %>%
  mutate(logct=log(ct))
plot(logct~age,data=d,pch=19)

## With 9-13
cc1 <- catchCurve(ct~age,data=d,ages2use=9:13)
sum1 <- cbind(summary(cc1),confint(cc1))
plot(cc1)
cc1w <- catchCurve(ct~age,data=d,ages2use=9:13,weighted=TRUE)
sum1w <- cbind(summary(cc1w),confint(cc1w))
cr1 <- chapmanRobson(ct~age,data=d,ages2use=9:13)
sum1cr <- cbind(summary(cr1),confint(cr1))
plot(cr1)

## With 8-13
cc2 <- catchCurve(ct~age,data=d,ages2use=8:13)
sum2 <- cbind(summary(cc2),confint(cc2))
plot(cc2)
cc2w <- catchCurve(ct~age,data=d,ages2use=8:13,weighted=TRUE)
sum2w <- cbind(summary(cc2w),confint(cc2w))
cr2 <- chapmanRobson(ct~age,data=d,ages2use=8:13)
sum2cr <- cbind(summary(cr2),confint(cr2))
plot(cr2)

## With 9-14
cc3 <- catchCurve(ct~age,data=d,ages2use=9:14)
sum3 <- cbind(summary(cc3),confint(cc3))
plot(cc3)
cc3w <- catchCurve(ct~age,data=d,ages2use=9:14,weighted=TRUE)
sum3w <- cbind(summary(cc3w),confint(cc3w))
cr3 <- chapmanRobson(ct~age,data=d,ages2use=9:14)
sum3cr <- cbind(summary(cr3),confint(cr3))
plot(cr3)

## With 8-14
cc4 <- catchCurve(ct~age,data=d,ages2use=8:14)
sum4 <- cbind(summary(cc4),confint(cc4))
plot(cc4)
cc4w <- catchCurve(ct~age,data=d,ages2use=8:14,weighted=TRUE)
sum4w <- cbind(summary(cc4w),confint(cc4w))
cr4 <- chapmanRobson(ct~age,data=d,ages2use=8:14)
sum4cr <- cbind(summary(cr4),confint(cr4))
plot(cr4)
```

#### With 9-13
1. This is a longitudinal catch-curve.
1. Ages 9-13.
1. XXX
    1. Z=`r sum1w["Z","Estimate"]` (CI: `r sum1w["Z","95% LCI"]`,`r sum1w["Z","95% UCI"]`)
    1. A=`r sum1w["A","Estimate"]` (CI: `r sum1w["A","95% LCI"]`,`r sum1w["A","95% UCI"]`)
    1. S=`r 100-sum1w["A","Estimate"]` (CI: `r 100-sum1w["A","95% UCI"]`,`r 100-sum1w["A","95% LCI"]`)
1. XXX
    1. Z=`r sum1cr["Z","Estimate"]` (CI: `r sum1cr["Z","95% LCI"]`,`r sum1cr["Z","95% UCI"]`)
    1. A=`r 100-sum1cr["S","Estimate"]` (CI: `r 100-sum1cr["S","95% UCI"]`,`r 100-sum1cr["S","95% LCI"]`)
    1. S=`r sum1cr["S","Estimate"]` (CI: `r sum1cr["S","95% LCI"]`,`r sum1cr["S","95% LCI"]`)
    1. A=1-e<sup>-Z</sup>


#### With 8-13
1. This is a longitudinal catch-curve.
1. Ages 8-13.
1. XXX
    1. Z=`r sum2w["Z","Estimate"]` (CI: `r sum2w["Z","95% LCI"]`,`r sum2w["Z","95% UCI"]`)
    1. A=`r sum2w["A","Estimate"]` (CI: `r sum2w["A","95% LCI"]`,`r sum2w["A","95% UCI"]`)
    1. S=`r 100-sum2w["A","Estimate"]` (CI: `r 100-sum2w["A","95% UCI"]`,`r 100-sum2w["A","95% LCI"]`)
1. XXX
    1. Z=`r sum2cr["Z","Estimate"]` (CI: `r sum2cr["Z","95% LCI"]`,`r sum2cr["Z","95% UCI"]`)
    1. A=`r 100-sum2cr["S","Estimate"]` (CI: `r 100-sum2cr["S","95% UCI"]`,`r 100-sum2cr["S","95% LCI"]`)
    1. S=`r sum2cr["S","Estimate"]` (CI: `r sum2cr["S","95% LCI"]`,`r sum2cr["S","95% LCI"]`)
1. A=1-e<sup>-Z</sup>


#### With 9-14
1. This is a longitudinal catch-curve.
1. Ages 9-14.
1. XXX
    1. Z=`r sum3w["Z","Estimate"]` (CI: `r sum3w["Z","95% LCI"]`,`r sum3w["Z","95% UCI"]`)
    1. A=`r sum3w["A","Estimate"]` (CI: `r sum3w["A","95% LCI"]`,`r sum3w["A","95% UCI"]`)
    1. S=`r 100-sum3w["A","Estimate"]` (CI: `r 100-sum3w["A","95% UCI"]`,`r 100-sum3w["A","95% LCI"]`)
1. XXX
    1. Z=`r sum3cr["Z","Estimate"]` (CI: `r sum3cr["Z","95% LCI"]`,`r sum3cr["Z","95% UCI"]`)
    1. A=`r 100-sum3cr["S","Estimate"]` (CI: `r 100-sum3cr["S","95% UCI"]`,`r 100-sum3cr["S","95% LCI"]`)
    1. S=`r sum3cr["S","Estimate"]` (CI: `r sum3cr["S","95% LCI"]`,`r sum3cr["S","95% LCI"]`)

1. A=1-e<sup>-Z</sup>



#### With 8-14
1. This is a longitudinal catch-curve.
1. Ages 8-14.
1. XXX
    1. Z=`r sum4w["Z","Estimate"]` (CI: `r sum4w["Z","95% LCI"]`,`r sum4w["Z","95% UCI"]`)
    1. A=`r sum4w["A","Estimate"]` (CI: `r sum4w["A","95% LCI"]`,`r sum4w["A","95% UCI"]`)
    1. S=`r 100-sum4w["A","Estimate"]` (CI: `r 100-sum4w["A","95% UCI"]`,`r 100-sum4w["A","95% LCI"]`)
1. XXX
    1. Z=`r sum4cr["Z","Estimate"]` (CI: `r sum4cr["Z","95% LCI"]`,`r sum4cr["Z","95% UCI"]`)
    1. A=`r 100-sum4cr["S","Estimate"]` (CI: `r 100-sum4cr["S","95% UCI"]`,`r 100-sum4cr["S","95% LCI"]`)
    1. S=`r sum4cr["S","Estimate"]` (CI: `r sum4cr["S","95% LCI"]`,`r sum4cr["S","95% LCI"]`)
1. A=1-e<sup>-Z</sup>

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<KBLT>>
```

----

## Stannard Rock Lake Trout
```{r echo=FALSE, results='hide', fig.show='hide',warning=FALSE}
d <- data.frame(age=6:14,
                yr63=c(129,339,331,192, 70, 16,0.5,0.5,0.5),
                yc63=c( 45, NA, NA, NA,491,163,117, 18,  7),
                yc54=c( NA, NA, NA,192, 63, 39, 5,0.5, 0.5),
                yc67=c(101,733,1517,1560,1031,445,300,NA,NA)) %>%
  mutate(log54=log(yc54),log67=log(yc67))
d

## 1963 Capture year
cc63 <- catchCurve(yr63~age,data=d,ages2use=8:11,weighted=TRUE)
sum63 <- cbind(summary(cc63),confint(cc63))
plot(cc63)
cr63 <- chapmanRobson(yr63~age,data=d,ages2use=8:11)
sum63cr <- cbind(summary(cr63),confint(cr63))

## 1963 Year-Class
cc63yc <- catchCurve(yc63~age,data=d,ages2use=10:14,weighted=TRUE)
sum63yc <- cbind(summary(cc63yc),confint(cc63yc))
plot(cc63yc)
cr63yc <- chapmanRobson(yc63~age,data=d,ages2use=10:14)
sum63yccr <- cbind(summary(cr63yc),confint(cr63yc))

## 1954 Year-Class
cc54yc <- catchCurve(yc54~age,data=d,ages2use=9:14,weighted=TRUE)
sum54yc <- cbind(summary(cc54yc),confint(cc54yc))
plot(cc54yc)
cr54yc <- chapmanRobson(yc54~age,data=d,ages2use=9:14)
sum54yccr <- cbind(summary(cr54yc),confint(cr54yc))

## 1967 Year-Class
cc67yc <- catchCurve(yc67~age,data=d,ages2use=9:12,weighted=TRUE)
sum67yc <- cbind(summary(cc67yc),confint(cc67yc))
plot(cc67yc)
cr67yc <- chapmanRobson(yc67~age,data=d,ages2use=9:12)
sum67yccr <- cbind(summary(cr67yc),confint(cr67yc))

## 1954 to 1967 Comparison
### Using ages 9-14
d2 <- stack(d,select=c(yc54,yc67))
names(d2) <- c("ct","grp")
d2 %<>% mutate(age=c(6:14,6:14),logct=log(ct)) %>%
  filterD(age>=9,age<=14)

lm1 <- lm(logct~age*grp,data=d2)
anova(lm1)
fitPlot(lm1)

### Using ages 9-12
d3 <- filterD(d2,age<=12)
lm2 <- lm(logct~age*grp,data=d3)
anova(lm2)
fitPlot(lm2)
```

1. 1956 year-class.
1. Shown below.
```{r echo=FALSE}
d[,1:5]
```
1. 
    1. 1963 Capture Year
        1. Cross-sectional
        1. Ages 8-11
        1. Z=`r sum63["Z","Estimate"]` (CI: `r sum63["Z","95% LCI"]`,`r sum63["Z","95% UCI"]`)
        1. A=`r sum63["A","Estimate"]`
        1. Z=`r sum63cr["Z","Estimate"]` (CI: `r sum63cr["Z","95% LCI"]`,`r sum63cr["Z","95% UCI"]`)
        1. A=`r 100-sum63cr["S","Estimate"]`
1. 
    1. 1963 Year-Class
        1. Longitudinal
        1. Ages 9-12
        1. Z=`r sum63yc["Z","Estimate"]` (CI: `r sum63yc["Z","95% LCI"]`,`r sum63yc["Z","95% UCI"]`)
        1. A=`r sum63yc["A","Estimate"]`
        1. Z=`r sum63yccr["Z","Estimate"]` (CI: `r sum63yccr["Z","95% LCI"]`,`r sum63yccr["Z","95% UCI"]`)
        1. A=`r 100-sum63yccr["S","Estimate"]`
1. 
    1. 1954 Year-Class
        1. Longitudinal
        1. Ages 9-12
        1. Z=`r sum54yc["Z","Estimate"]` (CI: `r sum54yc["Z","95% LCI"]`,`r sum54yc["Z","95% UCI"]`)
        1. A=`r sum54yc["A","Estimate"]`
        1. Z=`r sum54yccr["Z","Estimate"]` (CI: `r sum54yccr["Z","95% LCI"]`,`r sum54yccr["Z","95% UCI"]`)
        1. A=`r 100-sum54yccr["S","Estimate"]`
1. 
    1. 1967 Year-Class
        1. Longitudinal
        1. Ages 9-12
        1. Z=`r sum67yc["Z","Estimate"]` (CI: `r sum67yc["Z","95% LCI"]`,`r sum67yc["Z","95% UCI"]`)
        1. A=`r sum67yc["A","Estimate"]`
        1. Z=`r sum67yccr["Z","Estimate"]` (CI: `r sum67yccr["Z","95% LCI"]`,`r sum67yccr["Z","95% UCI"]`)
        1. A=`r 100-sum67yccr["S","Estimate"]`

