---
layout: page
title: Homework Key
subtitle: Depletion
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/CE/")
source("../../rhelpers/rhelpers.R")
modHTML("KEY_Depletion_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(FSA)
library(dplyr)
source("../../rhelpers/knitr_setup.R")
```

----

## Lake Guntersville Largemouth Bass
```{r LGLMB, echo=FALSE, results='hide', message=FALSE}
d <- data.frame(ct=c(23,12,13,14,9,7),
                ft=rep(1,6)) %>%
  mutate(CPE=ct/ft)
depl1 <- depletion(d$ct,d$ft,method="Leslie")
summary(depl1)
confint(depl1)
```
```{r echo=FALSE, results='hide'}
sum3 <- summary(depl1)
Nhat <- sum3["No","Estimate"]
q <- sum3["q","Estimate"]
ci3 <- confint(depl1)
```

1. A total of `r d$CPE[1]` fish were caught per hour in the first sample.
1. The catchability coefficient (q) is estimated to be `r formatC(q,format="f",digits=3)` with a 95% confidence interval from `r formatC(ci3["q","95% LCI"],format="f",digits=3)` to `r formatC(ci3["q","95% UCI"],format="f",digits=3)`. Thus, it is estimated that `r formatC(q*100,format="f",digits=1)`% of the Largemouth Bass (*Micropterus salmoides*) in the enclosures would be captured with one hour (i.e., one unit) of electrofishing effort. This value is the same as that reported in Maceina *et al.* (1995).
1. The initial population size (N<sub>0</sub>) is estimated to be `r formatC(Nhat,format="f",digits=0)` with a 95% confidence interval from `r formatC(ci3["No","95% LCI"],format="f",digits=0)` to `r formatC(ci3["No","95% UCI"],format="f",digits=0)`. This value is the same as that reported in Maceina *et al.* (1995).
1. One would expect to catch `r formatC(q,format="f",digits=3)`&times;`r formatC(Nhat,format="f",digits=1)`&times;1 = `r formatC(q*Nhat,format="f",digits=1)` fish in the first hour of sampling.
1. To compute the expected catch for the second sample, one must first reduce the original population size by the number of fish removed with the first sample (i.e., N<sub>0</sub>-N<sub>0</sub>&times;q&times;1) and then multiply that times q&times;1. Thus, the expected catch for the second sample is (`r formatC(Nhat,format="f",digits=1)`-`r formatC(q*Nhat,format="f",digits=1)`)&times;`r formatC(q,format="f",digits=3)`&times;1 = `r formatC(q*Nhat-q^2*Nhat,format="f",digits=1)` fish.

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<LGLMB>>
```

&nbsp;

## Benthic Fish in a Headwater Stream
```{r BH, echo=FALSE, results='hide', message=FALSE}
ct <- c(138,55,48)
rem1 <- removal(ct)
summary(rem1)
confint(rem1)
M1 <- removal(ct,method="Moran")
S1 <- removal(ct,method="Schnute")
teststat <- 2*(M1$min.nlogLH-S1$min.nlogLH)
pvalue <- pchisq(teststat,df=1,lower.tail=FALSE)
```
```{r echo=FALSE, results='hide'}
sum4 <- summary(rem1)
Nhat <- sum4["No","Estimate"]
p <- sum4["p","Estimate"]
ci4 <- confint(rem1)
n <- sum(ct)
```

1. The probability of capture (p) is estimated to be `r formatC(p,format="f",digits=3)` with a 95% confidence interval from `r formatC(ci4["p","95% LCI"],format="f",digits=3)` to `r formatC(ci4["p","95% UCI"],format="f",digits=3)`.
1. The initial population size (N<sub>0</sub>) is estimated to be `r formatC(Nhat,format="f",digits=0)` with a 95% confidence interval from `r formatC(ci4["No","95% LCI"],format="f",digits=0)` to `r formatC(ci4["No","95% UCI"],format="f",digits=0)`. This value is the same as that reported in Pritchard *et al.* (2021).
1. Capture efficiency is the percentage of the estimated number of Bullhead at the site (`r formatC(Nhat,format="f",digits=0)`) that were captured in the three electrofishing passes (`r n`) -- i.e., $100\frac{`r n`}{`r formatC(Nhat,format="f",digits=0)`}$=`r formatC(100*n/Nhat,format="f",digits=0)`%. This is the same value as reported in Pritchard *et al.* (2021).
1. There is sufficient evidence to conclude that the probability of capture differed after the first of the three electrofishing passes (`r kPvalue(pvalue,latex=FALSE)`).
