---
layout: page
title: Homework Key
subtitle: Mark-Recapture
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCNRS349/modules/MarkRecap/")
source("../../rhelpers/rhelpers.R")
modHTML("HW_Key")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(FSA)
library(FSAdata)
source("../../rhelpers/knitr_setup.R")
set.seed(13797834)
```

----

## Ugly River Brown Trout
```{r UglyMR, echo=FALSE, results='hide', message=FALSE}
mr1 <- mrClosed(43,123,16,method="Chapman")
summary(mr1)
confint(mr1,verbose=TRUE)
```
```{r echo=FALSE, results='hide'}
ci <- confint(mr1)
```

1. M=43, n=123, m=16, N is unknown
1. The estimated aboundance of Brown Trout (*Salmo trutta*) in this section of the Ugly River is `r formatC(mr1$N,format="f",digits=0)` fish with a 95% confidence interval from `r formatC(ci[1,"95% LCI"],format="f",digits=0)` to `r formatC(ci[1,"95% UCI"],format="f",digits=0)`.
1. The binomial distribution was used to construct the confidence intervals because more than 10% of the fish in the second sample were recaptured (i.e., `r formatC(mr1$m/mr1$n*100,format="f",digits=1)`).

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<UglyMR>>
```

----

## Upper Niagara Springs Pond Rainbow Trout
```{r UNSPMR, echo=FALSE, results='hide', message=FALSE}
data(RBTroutUNSP)
ch1 <- capHistSum(RBTroutUNSP)
mr2 <- mrClosed(ch1,method="Chapman")
summary(mr2)
confint(mr2,verbose=TRUE)
```
```{r echo=FALSE, results='hide'}
ci2 <- confint(mr2)
```

1. A total of `r ch1$caphist[["10"]]` fish were captured in the first but not the second sample, `r ch1$caphist[["01"]]` fish were captured in the second but not the first sample, and `r ch1$caphist[["11"]]` fish were captured in both samples.
1. M=`r ch1$sum[["M"]]`, n=`r ch1$sum[["n"]]`, m=`r ch1$sum[["m"]]`
1. The estimated aboundance of Rainbow Trout (*Oncorhynchus mykss*) greater than 100 mm in Upper Niagara Springs Pond is `r formatC(mr2$N,format="f",digits=0)` fish with a 95% confidence interval from `r formatC(ci2[1,"95% LCI"],format="f",digits=0)` to `r formatC(ci2[1,"95% UCI"],format="f",digits=0)`.
1. The Poisson distribution was used to construct the confidence intervals because `r mr2$m` (<50) fish, but not greater than 10% of the fish (i.e., `r formatC(mr2$m/mr2$n*100,format="f",digits=1)`), in the second sample were recaptured.

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<UNSPMR>>
```

----

## Wisconsin YOY Walleye
```{r WaeMR, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
mr3 <- mrClosed(n=c(321,412,178,415,367),
                m=c( NA, 45, 55, 93,113),
                R=c(321,412,178,415, NA),method="Schnabel")
summary(mr3)
confint(mr3,verbose=TRUE)
```
```{r echo=FALSE, results='hide'}
ci3 <- confint(mr3)
```

1. The estimated aboundance of YOY Walleye (*Sander vitreus*) in Pine Lake is `r formatC(mr3$N,format="f",digits=0)` fish with a 95% confidence interval from `r formatC(ci3[1,"95% LCI"],format="f",digits=0)` to `r formatC(ci3[1,"95% UCI"],format="f",digits=0)`.
1. The number of recaptured fish in the third sample appears too large (see Figure below) and suggests a potential problem with the assumptions for the Schnabel method, likely an increased catchability on that sample.

```{r echo=FALSE, fig.width=3.5, fig.height=3.5}
plot(mr3)
```


#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<WaeMR>>
```

